Hybrid platforms and bargaining power

### 2024-06-20 src/paper/application.tex
Some more appendix rearrangement

In the main text, I use a simpler version of this profit sharing rule.
I assume that $w_P(s) = w_F(s) \equiv 1$.
As \cref{sec:cooperative_game} demonstrates, this corresponds to players. Getting their Shapley values in the end.
Together with the assumptions on the demand system, one can even derive a closed form expression for the platform's profit share (\cref{prop:platform_profits_bargaining}).
The proof if this proposition is given below
\begin{proof}[Proof of \cref{prop:platform_profits_bargaining}]
    Simply integrate the total industry profit function with respect to the mass of fringe entrants obtain the platform's share of the pie:
    \begin{align*}
        \pi^t_P &= \int_0^1 f(N_P, sN_F) \ds \\
        &= \mu \int_0^1 \frac{N_P V_P + s N_F V_F}{N_P V_P + s N_F V_F + 1} \ds \\
        &= \mu \left[ 1 - \frac{\log \left(1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} \right].
    \end{align*}
    The fringe's share is just the remainder,
    \begin{align*}
        f(N_P, N_F) - \int_0^1 f(N_P, sN_F) \ds = \mu \left[ \frac{\log \left( 1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} - \frac{1}{N_P V_P + N_F V_F + 1} \right]
    \end{align*}
    Finally, remember that the investment cost of the fringe is fixed at the bargaining stage, and is therefore not included in the bargaining outcome.
    Therefore, the total profits of the complete fringe are
    \begin{align*}
        \pi^t_P = \mu \left[ \frac{\log \left( 1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} - \frac{1}{N_P V_P + N_F V_F + 1} \right] - I_F N_F.
    \end{align*}
\end{proof}

@@

\subsection{Results}
\label{sec:more_general_results}

This section presents a number of results that can be derived even in this rather abstract setting.
Let us start with a technical, but very useful result.
\begin{proposition}
    \label{prop:unique_equilibrium_more_general}
    Under the conditions in \cref{ass:free_entry,ass:single_crossing}, the equilibrium is unique if it exists.
\end{proposition}
\begin{proof}[Proof of \cref{prop:unique_equilibrium_more_general}]
    \Cref{lem:slope_at_eq} states that for any positive $N_F^*$ for which $\pi_F(N_P, N_F^*) - N_F I_F$, the partial derivative with respect to the number of fringe firms is negative.
    Now assume by contradiction that $\exists 0 < N_F^* < N_F^{**}$ such that $\pi_F(N_P, N_F^*) = I_F N_F^*$ and $\pi_F(N_P, N_F^{**}) = I_F N_F^{**}$.
    But then the mean value theorem implies that there is a $\bar{N}_F \in (N_F^*, N_F^{**})$ such that $\partial_F \pi_F(N_P, \bar{N}_F) = I_F$.
    This is a contradiction, as $\partial_F \pi_F(N_P, N_F) < I_F$ for all $N_F^* < N_F$.
\end{proof}

@@

The following two propositions are partial equilibrium results: consider the number of fringe entrants $N_F$ as fixed.
The first statement claims that the platform's profits are increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Let \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing} hold.
    Further assume that $f$ is continuously differentiable with respect to $N_P$ and also twice differentiable.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0.
    \end{align*}
\end{proposition}
\begin{proof}[Proof of \cref{prop:share_of_platform}]
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F, N_C)}{\partial N_P} &= N_C \frac{\partial}{\partial N_P} \int_0^1 w_P(s) f(N_P, N_F) \ds \\
        &= \int_0^1 w_P(s) \underbrace{\frac{\partial f(N_P, sN_F)}{\partial N_P}}_{> 0} \ds > 0
    \end{align*}
    for any non-negative, continuous $w_P(s)$.
\end{proof}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the slice of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is only guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.\footnote{
    In fact, one can show that this is the case if the cross-derivatives of $f$ are negative.
}
\cref{fig:increase_N_P_platform} shows an example of this situation.

@@

Finally, let us turn to equilibrium entry as a function of the platform's product variety.
The previous proposition implies an almost immediate corollary regarding the equilibrium number of fringe entrants.
\begin{corollary}
    \label{cor:fringe_entry}
    Let \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing,ass:free_entry,ass:single_crossing} hold.
    Furthermore, assume that $f, w$ are twice continuously differentiable.
    Let $N_F^*$ denote the equilibrium number of fringe firms.
    Let us also assume that $N_F^* > 0$, and that
    \begin{align*}
        \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0 \quad \forall n_F \leq N_F \\
        \frac{\partial^2 f(n_P, n_F)}{\partial n_F^2} < 0 \quad \forall n_F \leq N_F.
    \end{align*}
    Then the equilibrium number of fringe firms is also differentiable and
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < 0.
    \end{align*}
\end{corollary}
\begin{proof}[Proof of \cref{cor:fringe_entry}]
    \Cref{prop:unique_equilibrium} establishes that if an equilibrium with $N_F > 0$ exists, it is unique.
    Use the implicit function theorem on the equation from \cref{ass:free_entry} to obtain
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} = \frac{\frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P}}{I_F - \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F}}
    \end{align*}
    The derivative exists if the above expression is well-defined, i.e. $\frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} \neq I_F$.
    Remember from \cref{prop:share_of_fringe} that the numerator of this expression is negative under the condition $\frac{\partial^2 f(N_P, N_F, N_C)}{\partial N_P \partial N_F}$.
    Now all we need to show to conclude the proof is that
    \begin{align*}
        \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} < I_F,
    \end{align*}
    which is the statement of \cref{lem:slope_at_eq}.
\end{proof}

@@

\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing,ass:free_entry,ass:single_crossing,ass:additive_profit} hold.
    Then
    \begin{align*}
        \frac{\partial N_F^*}{\partial N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
    Furthermore,
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_P} < 0.
    \end{align*}
\end{proposition}
\begin{proof}[Proof of \cref{prop:aggregate_size_additive}]
    Use the implicit function theorem to get the equilibrium number of fringe firms as a function of the platform's product variety:
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} = \frac{\frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P}}{I_F - \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F}}.
    \end{align*}
    Remember, that from \cref{lem:slope_at_eq}, and the assumption on the investment cost function we have that $\frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} < I_F$.
    Then, substitute $f(N_P, N_F) = g(\alpha N_P + \beta N_P)$ into $\pi_F$ to get the following fringe profit function:
    \begin{align*}
        \pi_F(N_P, N_F, N_C) = \beta N_C N_F \int_0^1 w_F(s) g'(\alpha N_P + s \beta N_F) \ds.
    \end{align*}
    Differentiate it with respect to $N_P$ and $N_F$ and substitute into the first expression to obtain
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} = \frac{
            \alpha \beta N_C N_P \int_0^1 w_F(s) g''(\alpha N_P + s \beta N_F) \ds
        }{
            \underbracket{I'_F(N_F) - \beta N_C \int_0^1 w_F(s) g'(\alpha N_P + s \beta N_F) \ds}_{(1)} - \underbracket{\beta^2 N_C N_F \int_0^1 w_F(s) s g''(\alpha N_P + s \beta N_F) \ds}_{(2)}.
        }
    \end{align*}
    From \cref{lem:slope_at_eq}, we know that the denominator of this expression is positive.
    Furthermore, the concavity of $g$ implies that expression (2) is also positive.
    Finally, from \cref{prop:share_of_fringe}, we have that the whole expression must be negative.

@@

    Putting it all together, we have that
    \begin{align*}
        \frac{\partial N_F}{\partial N_P} \leq \frac{\alpha \beta N_C N_P \int_0^1 w_F(s) g''(\alpha N_P + s \beta N_F) \ds}{\beta^2 N_C N_F \int_0^1 w_F(s) g''(\alpha N_P + s \beta N_F) \ds} = -\frac{\alpha}{\beta},
    \end{align*}
    which is the first statement of the proposition.

    To prove the second part of the proposition, differentiate total profits with respect to $N_P$:
    \begin{align*}
        \frac{\partial \Pi(N_P, N_F^*(N_P), N_C)}{\partial N_F} &= \frac{\partial}{\partial N_F} N_C g(\alpha N_P + \beta N_F^*(N_P)) \\
        &= \underbrace{N_C g'(\alpha N_P + \beta N_F^*(N_P))}_{> 0} \left[ \alpha + \beta \frac{\partial N_F^*(N_P)}{\partial N_P} \right] \\
        &< N_C g'(\alpha N_P + \beta N_F^*(N_P)) \left[ \alpha + \beta \left( -\frac{\alpha}{\beta} \right) \right] \\
        &= 0.
    \end{align*}
\end{proof}

@@

>>>>

In the main text, I use a simpler version of this profit sharing rule.
I assume that $w_P(s) = w_F(s) \equiv 1$.
As \cref{sec:cooperative_game} demonstrates, this corresponds to players. Getting their Shapley values in the end.
Together with the assumptions on the demand system, one can even derive a closed form expression for the platform's profit share (\cref{prop:platform_profits_bargaining}).
The proof if this proposition is given below.
\begin{proof}[Proof of \cref{prop:platform_profits_bargaining}]
    Simply integrate the total industry profit function with respect to the mass of fringe entrants obtain the platform's share of the pie:
    \begin{align*}
        \pi^t_P &= \int_0^1 f(N_P, sN_F) \ds \\
        &= \mu \int_0^1 \frac{N_P V_P + s N_F V_F}{N_P V_P + s N_F V_F + 1} \ds \\
        &= \mu \left[ 1 - \frac{\log \left(1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} \right].
    \end{align*}
    The fringe's share is just the remainder,
    \begin{align*}
        f(N_P, N_F) - \int_0^1 f(N_P, sN_F) \ds = \mu \left[ \frac{\log \left( 1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} - \frac{1}{N_P V_P + N_F V_F + 1} \right]
    \end{align*}
    Finally, remember that the investment cost of the fringe is fixed at the bargaining stage, and is therefore not included in the bargaining outcome.
    Therefore, the total profits of the complete fringe are
    \begin{align*}
        \pi^t_P = \mu \left[ \frac{\log \left( 1 + \frac{N_F V_F}{N_P V_P + 1} \right)}{N_F V_F} - \frac{1}{N_P V_P + N_F V_F + 1} \right] - I_F N_F.
    \end{align*}
\end{proof}

@@

With \cref{ass:single_crossing} in place, uniqueness of the equilibrium can be established.
\begin{proposition}
    \label{prop:unique_equilibrium_more_general}
    Under the conditions in \cref{ass:free_entry,ass:single_crossing}, the equilibrium is unique if it exists.
\end{proposition}
\begin{proof}[Proof of \cref{prop:unique_equilibrium_more_general}]
    \Cref{lem:slope_at_eq} states that for any positive $N_F^*$ for which $\pi_F(N_P, N_F^*) - N_F I_F$, the partial derivative with respect to the number of fringe firms is negative.
    Now assume by contradiction that $\exists 0 < N_F^* < N_F^{**}$ such that $\pi_F(N_P, N_F^*) = I_F N_F^*$ and $\pi_F(N_P, N_F^{**}) = I_F N_F^{**}$.
    But then the mean value theorem implies that there is a $\bar{N}_F \in (N_F^*, N_F^{**})$ such that $\partial_F \pi_F(N_P, \bar{N}_F) = I_F$.
    This is a contradiction, as $\partial_F \pi_F(N_P, N_F) < I_F$ for all $N_F^* < N_F$.
\end{proof}

@@

\subsection{Comparative statics}
\label{sec:more_general_results}

This section presents a number of comparative statics that can be derived even in this rather abstract setting.
It contains three sets of results: (1) participants' profits in a partial equilibrium setting, where the number of fringe firms is taken as fixed, (2) equilibrium entry as a function of the platform's product variety, and (3) the platform's profits in general equlibrium.
Throughout the paper, I use $\frac{\partial X}{\partial N_P}$ to denote partial equilibrium results, while $\frac{\dd X}{\dd N_P} \coloneq \frac{\partial X(N_P, N_F(N_P))}{\partial N_P}$ indicates general equilibrium results, where fringe entry is endogeneous.

\paragraph{Profits -- partial equilibrium}
The following two propositions are partial equilibrium results: consider the number of fringe entrants $N_F$ as fixed.
The first statement claims that the platform's profits are increasing in its own product variety.
\begin{proposition}
    \label{prop:share_of_platform}
    Let \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing} hold.
    Further assume that $f$ is continuously differentiable with respect to $N_P$ and also twice differentiable.
    Let $N_F \geq 0$.
    Then $\pi_P$ is also differentiable and
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F)}{\partial N_P} > 0.
    \end{align*}
\end{proposition}
\begin{proof}[Proof of \cref{prop:share_of_platform}]
    $f(N_P, N_F)$ is continuously differentiable in $N_P$, therefore the Leibniz rule can be applied to obtain
    \begin{align*}
        \frac{\partial \pi_P(N_P, N_F, N_C)}{\partial N_P} &= N_C \frac{\partial}{\partial N_P} \int_0^1 w_P(s) f(N_P, N_F) \ds \\
        &= \int_0^1 w_P(s) \underbrace{\frac{\partial f(N_P, sN_F)}{\partial N_P}}_{> 0} \ds > 0
    \end{align*}
    for any non-negative, continuous $w_P(s)$.
\end{proof}
While it might seem obvious, let us still examine what this result does and does not mean.
First, remember that $f$ is increasing in both arguments, and $N_F$ is fixed.
Therefore, an increase in $N_P$ also increases the size of the pie the participants bargain over.
This result states that the slice of the pie that the platform gets increases in this case, too.
It does not mean, however, that the \emph{relative share} of the pie that the platform gets is also bigger -- increase is only guaranteed in absolute terms.
For example, it is possible that for the new, higher value of $N_P$, the complementarities between the fringe firms become stronger, and the platform's bargaining power decreases.\footnote{
    In fact, one can show that this is the case if the cross-derivatives of $f$ are negative.
}
\cref{fig:increase_N_P_platform} shows an example of this situation.

@@

\paragraph{Number of entrants -- general equilibrium}
Next, let us turn to equilibrium entry as a function of the platform's product variety.
The previous proposition implies an almost immediate corollary regarding the equilibrium number of fringe entrants.
\begin{corollary}
    \label{cor:fringe_entry}
    Let \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing,ass:free_entry,ass:single_crossing} hold.
    Furthermore, assume that $f, w$ are twice continuously differentiable.
    Let $N_F^*$ denote the equilibrium number of fringe firms.
    Let us also assume that $N_F^* > 0$, and that
    \begin{align*}
        \frac{\partial^2 f(n_P, n_F)}{\partial n_P \partial n_F} < 0 \quad \forall n_F \leq N_F \\
        \frac{\partial^2 f(n_P, n_F)}{\partial n_F^2} < 0 \quad \forall n_F \leq N_F.
    \end{align*}
    Then the equilibrium number of fringe firms is also differentiable and
    \begin{align*}
        \dd{\partial N_F}{\dd N_P} < 0.
    \end{align*}
\end{corollary}
\begin{proof}[Proof of \cref{cor:fringe_entry}]
    \Cref{prop:unique_equilibrium} establishes that if an equilibrium with $N_F > 0$ exists, it is unique.
    Use the implicit function theorem on the equation from \cref{ass:free_entry} to obtain
    \begin{align*}
        \frac{\dd N_F}{\dd N_P} = \frac{\frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P}}{I_F - \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F}}
    \end{align*}
    The derivative exists if the above expression is well-defined, i.e. $\frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} \neq I_F$.
    Remember from \cref{prop:share_of_fringe} that the numerator of this expression is negative under the condition $\frac{\partial^2 f(N_P, N_F, N_C)}{\partial N_P \partial N_F}$.
    Now all we need to show to conclude the proof is that
    \begin{align*}
        \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} < I_F,
    \end{align*}
    which is the statement of \cref{lem:slope_at_eq}.
\end{proof}

@@

\begin{proposition}
    \label{prop:aggregate_size_additive}
    Let \cref{ass:identical_fringe,ass:monotone_profits,ass:profit_sharing,ass:free_entry,ass:single_crossing,ass:additive_profit} hold.
    Then
    \begin{align*}
        \frac{\dd N_F}{\dd N_P} < -\frac{\alpha}{\beta}.
    \end{align*}
    Furthermore,
    \begin{align*}
        \frac{\dd \Pi(N_P, N_F, N_C)}{\dd N_P} < 0.
    \end{align*}
\end{proposition}
\begin{proof}[Proof of \cref{prop:aggregate_size_additive}]
    Use the implicit function theorem to get the equilibrium number of fringe firms as a function of the platform's product variety:
    \begin{align*}
        \frac{\dd N_F}{\dd N_P} = \frac{\frac{\partial \pi_F(N_P, N_F, N_C)}{\partial N_P}}{I_F - \frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F}}.
    \end{align*}
    Remember, that from \cref{lem:slope_at_eq}, and the assumption on the investment cost function we have that $\frac{\partial \pi_F (N_P, N_F, N_C)}{\partial N_F} < I_F$.
    Then, substitute $f(N_P, N_F) = g(\alpha N_P + \beta N_P)$ into $\pi_F$ to get the following fringe profit function:
    \begin{align*}
        \pi_F(N_P, N_F, N_C) = \beta N_C N_F \int_0^1 w_F(s) g'(\alpha N_P + s \beta N_F) \ds.
    \end{align*}
    Differentiate it with respect to $N_P$ and $N_F$ and substitute into the first expression to obtain
    \begin{align*}
        \frac{\dd N_F}{\dd N_P} = \frac{
            \alpha \beta N_C N_P \int_0^1 w_F(s) g''(\alpha N_P + s \beta N_F) \ds
        }{
            \underbracket{I'_F(N_F) - \beta N_C \int_0^1 w_F(s) g'(\alpha N_P + s \beta N_F) \ds}_{(1)} - \underbracket{\beta^2 N_C N_F \int_0^1 w_F(s) s g''(\alpha N_P + s \beta N_F) \ds}_{(2)}.
        }
    \end{align*}
    From \cref{lem:slope_at_eq}, we know that the denominator of this expression is positive.
    Furthermore, the concavity of $g$ implies that expression (2) is also positive.
    Finally, from \cref{prop:share_of_fringe}, we have that the whole expression must be negative.

@@

    Putting it all together, we have that
    \begin{align*}
        \frac{\dd N_F}{\dd N_P} \leq \frac{\alpha \beta N_C N_P \int_0^1 w_F(s) g''(\alpha N_P + s \beta N_F) \ds}{\beta^2 N_C N_F \int_0^1 w_F(s) g''(\alpha N_P + s \beta N_F) \ds} = -\frac{\alpha}{\beta},
    \end{align*}
    which is the first statement of the proposition.

    To prove the second part of the proposition, differentiate total profits with respect to $N_P$:
    \begin{align*}
        \frac{\dd \Pi(N_P, N_F, N_C)}{\dd N_P}{\dd} &= \frac{\partial \Pi(N_P, N_F(N_P), N_C)}{\partial N_P} \\
         &= \frac{\partial}{\partial N_P} N_C g(\alpha N_P + \beta N_F^*(N_P)) \\
        &= \underbrace{N_C g'(\alpha N_P + \beta N_F^*(N_P))}_{> 0} \left[ \alpha + \beta \frac{\partial N_F^*(N_P)}{\partial N_P} \right] \\
        &< N_C g'(\alpha N_P + \beta N_F^*(N_P)) \left[ \alpha + \beta \left( -\frac{\alpha}{\beta} \right) \right] \\
        &= 0.
    \end{align*}
\end{proof}

@@

\paragraph{Platform profits -- general equilibrium}
Finally, let us examine how the platform's product variety impacts its total profits, also taking into account its effect on the number of fringe entrants.
